Growth stresses form in most tree species, in Eucalyptus (and many other
species) these internal stresses can be of a considerable magnitude, especially
in reaction wood. When a stem or branch is cut, some of the stress is released
in the form of strain, resulting in a number of timber defects which reduce log
value.

It has been shown previously that these stresses are induced during secondary
cell wall formation, however the mechanism responsible is unknown. Two
hypotheses exist to explain the stress generation; 1) deposition of lignin
between the fibril aggregates causing the cell to expand or contract axially
depending on the orientation of the fibril (MFA), or 2) a mechanism resulting in
a length change of the cellulose itself causes cellular contraction or expansion
depending on the MFA. A combination of the two has also been proposed.

The proposed research will increase our understanding of the mechanisms behind
growth stress formation through the development of a mathematical model taking
into account a realistic cell wall supramolecular architecture to simulate
maturation. The geometric parameters of the cell wall molecular architecture and
chemical composition will be studied using x-ray diffraction, atomic force
microscopy and wet chemistry techniques. Experimentation will also be used to
investigate the role of the gelatinous (G) cell wall layer in the formation of
extreme tension wood.

The New Zealand Dryland Forest Initiative (NZDFI) is currently breeding
\textit{Eucalyptus bosistoana} as a high value timber crop alternative to
\textit{Pinus radiata}. One of the goals of the breeding trials is to select for
minimal growth stresses so that the timber does not lose value when harvesting
and milling due to the release of internal stresses. The proposed research will
select families at young ages which are producing the lowest possible growth
stresses improving the \textit{E. bosistoana} breeding stock.